Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-25 src/tex/ms.tex
Add comments

The two loss functions are chosen to give different 
behaviour during the optimization process.
The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
but is prone to be dominated by a single point in parameter space with a large mismatch.
Other points with smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such a loss function.
Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
by encouraging the waveform to improve at each training point at a similar rate.
The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
In this paper, we show the results of using both loss functions 
and examine the differences between them.

@@

\begin{figure}
	\script{loss.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/loss.pdf}
	\caption{Loss functions against the number of iterations.
	The vertical dotted line indicates minimum of the validation loss,
	where we stop the optimization and obtain a new set of waveform coefficients, $\lambda$.}
	\label{fig:loss}
\end{figure}

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the $q-\chi_{PN}$
	space. The blue stars represent the waveforms used to compute the training loss and the orange
	dots represent the waveforms used for validation.}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, nine of them are also used in the original IMRPhenomD calibration. The remaining two
	are SXS analogues to two of the BAM NR simulations used in the original calibration
	\citep{khan2016frequency}.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

\begin{figure}[t]
	\script{ps_q148_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
	\caption{Fractional mismatch change of the validation waveforms.
	Each quadrant is fitted independently with to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
	\label{fig:ps_q148_quadrant}
\end{figure}

@@

\kw{End here.}

% As discussed in Sec.~\ref{sec:method} our recalibration procedure is not exactly the same as the original
% calibration. For instance, we use a different set of NR waveforms, frequency
% ranges, etc. Nonetheless, since the decrease in mismatch is rather significant, we believe that
% the high-dimensional optimization procedure is primarily responsible to the improved performance.

>>>>

The two loss functions are chosen to give different 
behavior during the optimization process.
The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
but is prone to be dominated by a single point in parameter space with a large mismatch.
Other points with smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such a loss function.
Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
by encouraging the waveform to improve at each training point at a similar rate.
The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
In this paper, we show the results of using both loss functions 
and examine the differences between them.

@@

\begin{figure}
	\script{loss.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/loss.pdf}
	\caption{Loss functions against the number of iterations. The vertical
	dotted line indicates minimum of the validation loss, which is where we stop
	the optimization.}
	\label{fig:loss}
\end{figure}

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the
	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
	the training loss and the orange dots represent the waveforms used for
	validation. \kw{It is hard to see the blue stars on top of the orange. Play
	around with the size of the dots and change the colors. Also add the extra
	NR waveforms in table 2 with a different color but use the stars.}}}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, nine of them are also used in the original IMRPhenomD calibration. The remaining two
	are SXS analogues to two of the BAM NR simulations used in the original calibration
	\citep{khan2016frequency}. \kw{BE EXPLICIT ABOUT WHAT DO YOU MEAN THE REMAINING TWO!}}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

\begin{figure}[t]
	\script{ps_q148_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
	\caption{Fractional mismatch change of the validation waveforms.
	Each quadrant is fitted independently to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
	\label{fig:ps_q148_quadrant}
\end{figure}

@@
